---
title: 'Reading Guide: Test 2 - Chapter 15 Analysis of Biological Data 2nd ed.'
author: "brouwern@gmail.com"
date: "November 4, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CHAPTER 15: COMPARING MEANS OF MORE THAN TWO GROUPS

### 15.1 The Analysis of variance

#### Example 15.1 The knees who say night

You should be able to read an example like this and know that ANOVA needs to be used to analyze the data

#### Figure 15.1-1

You should be able to look at graph like this and know that ANOVA needs to be used to analyze the data


#### Hypotheses

You shoudl be able to relate the hypotheses in terms of mu and know what the null hypothesis and alt. hypothesees are.

#### ANOVA in a nutshell

[2017 - skipped]

#### ANOVA tables

[2017 - skipped]

#### Partitioning the sum of square (pg 464) 

[2017 - skipped]

#### Calculating the mean square (pg 466)

[2017 - skipped]

#### The Variance ratio, F (pg 467)

[2017 - skipped]

#### Variation explained R2

[2017 - skipped]

#### ANOVA with two groups

[2017 - skipped]

### 15.2 Assumptions and alternatives

Will cover this in test 3

#### Robustness of ANOVA (pg 470)

Will cover this in test 3

#### Data transformations (pg 471)

Will cover this in test 3

#### Nonparametric alternatives to ANOVA (pg 471)

Will cover this in test 3


```{r eval=FALSE, include=FALSE}
 # You should know that there are non-parametric alternatives to ANOVA. These involve converting the data to their ranks.

```


### 15.3 Planned comparisons

We covered this concept in detail

**Planned comparison**

#### Planned comparisons between means (pg 472-3)

We skipped the math

### 15.4 Unplanned comparisons

The method they focus on, the **Tukey-Kramer test**, is carried out using the TukeyHSD() function in R.  They discuss the **Bonferonni correction** in a seperate Interleaf.

### Example 15.4 Wood Wide Web

You should understand how the hypotheses are formulated and the Tukey test interpretted.  

### Assumptions (pg 477)

We will cover this on test 3.

#### [15.5 Fixed and Random effects ]

[Skip]

#### [15.6 ANOVA with randomly chosen groups - SKIP!]

[Skip]

### 15.7 Summary

Read appropriate sections.
